---
title: "Publication-Ready Tables in R"
subtitle: "A Practical Guide Using gtsummary and flextable"
author: "Muntasir Masum, PhD"
date: last-modified
date-format: "MMMM D, YYYY"

format:
  html:
    output-file: index.html
    theme: flatly
    toc: true
    toc-depth: 3
    toc-float: true
    toc-title: "Contents"
    code-fold: show
    code-tools: true
    number-sections: true
    smooth-scroll: true
    embed-resources: true
    fig-align: center
    df-print: paged
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
  cache: false

knitr:
  opts_chunk:
    comment: "#>"
---

```{r}
#| label: setup
#| include: false

# ── Packages ──────────────────────────────────────────────────────────────────
library(tidyverse)
library(labelled)
library(survey)
library(gtsummary)
library(gt)
library(flextable)

# ── Global gtsummary theme ────────────────────────────────────────────────────
options(gtsummary.use_ftExtra = TRUE)
set_gtsummary_theme(theme_gtsummary_compact(set_theme = TRUE))

# ── Load data ─────────────────────────────────────────────────────────────────
nhis <- readRDS("data/nhis_sample.rds")

# ── Survey design (full sample for descriptive tables) ───────────────────────
options(survey.lonely.psu = "adjust")

des <- svydesign(
  id      = ~psu,
  strata  = ~strata,
  weights = ~wgt,
  data    = nhis,
  nest    = TRUE
)

# ── Complete-case subset for regression ──────────────────────────────────────
nhis_cc <- nhis |>
  drop_na(poor_health, us_born, genx_mil, male, race_eth15,
          edu4, famincome, marital_stat2, smoker, alcohol,
          obese, hyperten, region2)

des_cc <- svydesign(
  id      = ~psu,
  strata  = ~strata,
  weights = ~wgt,
  data    = nhis_cc,
  nest    = TRUE
)
```

# Overview {.unnumbered}

Producing clean, publication-ready tables is one of the most important — and
most under-taught — skills in applied social science and public health
research. This tutorial walks through a complete workflow using two
complementary R packages:

| Package | Strength | Best for |
|---------|----------|----------|
| **gtsummary** | Automatic table building with survey support | Table 1, regression tables |
| **flextable** | Fine-grained formatting control | Word / .docx export |

All examples use a 20,000-observation random sample of the
[National Health Interview Survey (NHIS)](https://nhis.ipums.org/), linked to
the National Death Index through the NHIS Linked Mortality Files (LMF). The
analysis illustrates a common demography and public health workflow: describing
a sample, then modeling a binary health outcome with survey-weighted logistic
regression.

::: {.callout-tip}
## What you will learn

- How to build a complete Table 1 (descriptive statistics), with and without
  survey weights
- How to control variable type summaries (continuous vs. categorical)
- How to produce and format regression tables from `svyglm()` output
- Advanced gtsummary techniques: removing reference rows, significance stars,
  managing footnotes, `tbl_merge()`, and `tbl_stack()`
- How to export polished tables to Word via flextable
:::

# Data and Variables

The analytic sample contains `r nrow(nhis)` respondents drawn from NHIS waves
collected between 1997 and 2018, with variables spanning demographics,
socioeconomic status, health behaviors, and health outcomes.

```{r}
#| label: tbl-data-dictionary
#| echo: false
#| tbl-cap: "Variables in the Teaching Dataset"

tibble::tribble(
  ~Variable,        ~Label,
  "us_born",        "U.S.-born vs. Foreign-born",
  "genx_mil",       "Birth cohort (Generation X / Millennial)",
  "age_cat",        "Age group",
  "male",           "Sex",
  "race_eth15",     "Race/ethnicity (6 categories)",
  "edu4",           "Educational attainment (4 levels)",
  "famincome",      "Family income category",
  "marital_stat2",  "Marital status",
  "region2",        "U.S. Census region",
  "smoker",         "Current smoker",
  "alcohol",        "Heavy alcohol use",
  "obese",          "Obesity status",
  "hyperten",       "Hypertension diagnosis",
  "healthstatus",   "Self-rated health (Healthy / Poor)",
  "poor_health",    "Poor self-rated health (0 = Healthy, 1 = Poor)",
  "wgt",            "Mortality analysis weight (mortwtsa)",
  "strata",         "Survey stratum",
  "psu",            "Primary sampling unit"
) |>
  gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(13))
```

## Survey design {.unnumbered}

NHIS uses a complex multistage probability sample. Correct inference requires
declaring the survey design before any analysis:

```{r}
#| label: survey-design-demo
#| eval: false

library(survey)
options(survey.lonely.psu = "adjust")

des <- svydesign(
  id      = ~psu,       # primary sampling unit
  strata  = ~strata,    # stratification variable
  weights = ~wgt,       # mortality analysis weight (mortwtsa)
  data    = nhis,
  nest    = TRUE        # PSUs nested within strata
)
```

::: {.callout-important}
Always use `nest = TRUE` when PSU IDs are not unique across strata (which is
the case with NHIS). Setting `options(survey.lonely.psu = "adjust")` prevents
errors in strata with only one PSU.
:::

## Global gtsummary options {.unnumbered}

Two global options should be set once per session, before any table is built.
They apply automatically to every subsequent `tbl_summary()`,
`tbl_svysummary()`, and `tbl_regression()` call.

```{r}
#| label: gtsummary-options
#| eval: false

options(gtsummary.use_ftExtra = TRUE)
set_gtsummary_theme(theme_gtsummary_compact(set_theme = TRUE))
```

::: {.callout-tip}
## Why these two options matter

**`options(gtsummary.use_ftExtra = TRUE)`**

Activates the **ftExtra** backend for flextable rendering. By default,
gtsummary uses a basic text renderer when converting tables to flextable via
`as_flex_table()`. With ftExtra enabled, markdown in cell content (bold labels,
italic levels, superscripts, footnote symbols) is preserved faithfully in the
Word export. Without it, formatted text can appear as raw markdown syntax in
the `.docx` file.

**`set_gtsummary_theme(theme_gtsummary_compact(set_theme = TRUE))`**

Applies the **compact theme** globally. Compared to the default theme, compact
reduces row padding and font size, producing tables that fit comfortably on a
journal page. Using `set_gtsummary_theme()` as the outer wrapper is the
recommended pattern in gtsummary 2.x: it registers the theme with the package
engine so it persists across all tables in the session. Passing
`set_theme = TRUE` inside `theme_gtsummary_compact()` ensures the theme is
active immediately even if called in isolation.

Both lines are already active in this tutorial's hidden setup chunk.
:::

# Part 1: Descriptive Tables (Table 1)

Table 1 in a manuscript reports sample characteristics. The tabs below
progress from the simplest approach to a fully stratified, weighted table
ready for publication.

::: {.panel-tabset}

## Unweighted

`tbl_summary()` works directly on a data frame — no survey design needed.
Appropriate for convenience samples, or as a first pass before adding weights.

```{r}
#| label: tbl1-unweighted

nhis |>
  select(us_born, genx_mil, age_cat, male, race_eth15,
         edu4, famincome, healthstatus) |>
  tbl_summary(
    label = list(
      us_born      ~ "Nativity",
      genx_mil     ~ "Birth cohort",
      age_cat      ~ "Age group",
      male         ~ "Sex",
      race_eth15   ~ "Race/ethnicity",
      edu4         ~ "Education",
      famincome    ~ "Family income",
      healthstatus ~ "Self-rated health"
    ),
    missing = "no"
  ) |>
  bold_labels() |>
  modify_caption("**Table 1. Sample Characteristics (Unweighted)**") |>
  as_gt()
```

## Survey-Weighted

`tbl_svysummary()` accepts the `svydesign` object and weights all statistics
automatically. Proportions and means can differ substantially from the
unweighted version when survey weights correct for unequal selection
probabilities.

```{r}
#| label: tbl1-weighted

des |>
  tbl_svysummary(
    include = c(us_born, genx_mil, age_cat, male, race_eth15,
                edu4, famincome, healthstatus),
    label = list(
      us_born      ~ "Nativity",
      genx_mil     ~ "Birth cohort",
      age_cat      ~ "Age group",
      male         ~ "Sex",
      race_eth15   ~ "Race/ethnicity",
      edu4         ~ "Education",
      famincome    ~ "Family income",
      healthstatus ~ "Self-rated health"
    ),
    missing = "no"
  ) |>
  bold_labels() |>
  modify_caption("**Table 1. Sample Characteristics (Survey-Weighted)**") |>
  as_gt()
```

## Stratified by Nativity

Add `by =` to compare groups side by side. Chain `add_overall()` for a
total column, and `add_p()` for group-comparison p-values. Use
`modify_header()` with `{n}` (unweighted count) to show interpretable
sample sizes in column headers.

```{r}
#| label: tbl1-stratified

tbl1 <- des |>
  tbl_svysummary(
    by = us_born,
    include = c(genx_mil, age_cat, male, race_eth15,
                edu4, famincome, smoker, alcohol, obese,
                hyperten, healthstatus),
    label = list(
      genx_mil     ~ "Birth cohort",
      age_cat      ~ "Age group",
      male         ~ "Sex",
      race_eth15   ~ "Race/ethnicity",
      edu4         ~ "Education",
      famincome    ~ "Family income",
      smoker       ~ "Current smoker",
      alcohol      ~ "Heavy alcohol use",
      obese        ~ "Obese",
      hyperten     ~ "Hypertension",
      healthstatus ~ "Self-rated health"
    ),
    missing = "no"
  ) |>
  add_overall(last = FALSE) |>
  add_p(test.args = all_tests("svy.wilcox.test") ~ list(design = des)) |>
  bold_labels() |>
  italicize_levels() |>
  bold_p(t = 0.05) |>
  modify_header(
    stat_0 ~ "**Overall**\nn = {n_unweighted}",
    stat_1 ~ "**U.S.-Born**\nn = {n_unweighted}",
    stat_2 ~ "**Foreign-Born**\nn = {n_unweighted}"
  ) |>
  modify_caption("**Table 1. Sample Characteristics by Nativity**")

tbl1 |> as_gt()
```

::: {.callout-tip}
## Header placeholder guide for `tbl_svysummary`

| Placeholder | Returns | Use when |
|-------------|---------|----------|
| `{n_unweighted}` | Unweighted count per group | Column headers (most readable) |
| `{N_unweighted}` | Total unweighted N | Overall column header |
| `{n}` | Sum of weights per group | Reporting weighted N |
| `{N}` | Total sum of weights | Rarely useful in headers |

Both `{n}` and `{N}` return decimals in weighted surveys — use
`{n_unweighted}` for clean integer counts in column headers.
:::

:::

## Controlling Variable Type Summaries

By default, gtsummary guesses whether each variable is categorical or
continuous. You can override this with the `type =` argument, and customize
the displayed statistics with `statistic =`.

::: {.panel-tabset}

### Default behavior

```{r}
#| label: tbl-type-default

nhis |>
  select(age_cat, edu4, poor_health) |>
  tbl_summary(missing = "no") |>
  bold_labels() |>
  as_gt()
```

### Override: force continuous

Use `type = list(var ~ "continuous")` to display mean ± SD instead of
counts. Useful for variables that happen to be stored as factors but have
a meaningful numeric interpretation (e.g., an age index).

```{r}
#| label: tbl-type-continuous

nhis |>
  select(age_cat, edu4, poor_health) |>
  tbl_summary(
    type    = list(poor_health ~ "continuous"),
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) |>
  bold_labels() |>
  as_gt()
```

### Override: force categorical

Use `type = list(var ~ "categorical")` to show counts and percentages for a
numeric variable (e.g., an integer 0/1 outcome where you want both values
displayed explicitly).

```{r}
#| label: tbl-type-categorical

nhis |>
  select(age_cat, edu4, poor_health) |>
  tbl_summary(
    type    = list(poor_health ~ "categorical"),
    missing = "no"
  ) |>
  bold_labels() |>
  as_gt()
```

:::

::: {.callout-tip}
## Key gtsummary verbs for Table 1

| Function | Purpose |
|----------|---------|
| `tbl_svysummary(by = ...)` | Stratify columns by a grouping variable |
| `add_overall()` | Append an overall (unstratified) column |
| `add_p()` | Add p-values for group comparisons |
| `type = list(var ~ "continuous")` | Force a variable to display mean/SD |
| `type = list(var ~ "categorical")` | Force a variable to display counts/% |
| `statistic = list(...)` | Customize the displayed summary statistic |
| `bold_labels()` | Bold the variable name rows |
| `italicize_levels()` | Italicize the category rows |
| `bold_p(t = 0.05)` | Bold significant p-values |
| `modify_header()` | Rewrite any column header (use `{n}` for counts) |
| `modify_caption()` | Add or change table caption |
:::

# Part 2: Regression Tables

We model **poor self-rated health** (binary: 1 = Poor, 0 = Healthy) using
survey-weighted logistic regression via `svyglm()` with a `quasibinomial()`
family.

```{r}
#| label: fit-models

# Model A: Demographic only
m_A <- svyglm(
  poor_health ~ us_born + genx_mil + male + race_eth15,
  design = des_cc,
  family  = quasibinomial()
)

# Model B: + Socioeconomic status
m_B <- svyglm(
  poor_health ~ us_born + genx_mil + male + race_eth15 +
                edu4 + famincome,
  design = des_cc,
  family  = quasibinomial()
)

# Model C: Full model
m_C <- svyglm(
  poor_health ~ us_born + genx_mil + male + race_eth15 +
                edu4 + famincome + marital_stat2 +
                smoker + alcohol + obese + hyperten + region2,
  design = des_cc,
  family  = quasibinomial()
)
```

The four tabs below demonstrate progressively more refined formatting,
starting from the default output and building toward a publication-ready table.

::: {.panel-tabset}

## Basic Table

`tbl_regression()` converts any model object into a formatted table.
Set `exponentiate = TRUE` to display odds ratios instead of log-odds.

```{r}
#| label: tbl-regression-basic

tbl_regression(m_A, exponentiate = TRUE) |>
  bold_labels() |>
  as_gt()
```

## Remove Reference Rows

By default, gtsummary shows a reference row for each categorical predictor.
These rows add visual clutter. A single call to
`remove_row_type(type = "reference")` cleans them up.

::: {.panel-tabset}

### Before

```{r}
#| label: tbl-ref-before

tbl_regression(m_A, exponentiate = TRUE) |>
  bold_labels() |>
  italicize_levels() |>
  as_gt()
```

### After

```{r}
#| label: tbl-ref-after

tbl_regression(m_A, exponentiate = TRUE) |>
  remove_row_type(type = "reference") |>
  bold_labels() |>
  italicize_levels() |>
  as_gt()
```

:::

## Significance Stars

Some journals prefer the European style: coefficient (SE) with significance
stars rather than a confidence interval column.
Use `add_significance_stars()` with `conf.int = FALSE`.

```{r}
#| label: tbl-stars

tbl_regression(
  m_A,
  exponentiate = TRUE,
  conf.int     = FALSE
) |>
  add_significance_stars(
    hide_ci  = TRUE,
    hide_se  = FALSE,
    pattern  = "{estimate}{stars}"
  ) |>
  remove_row_type(type = "reference") |>
  bold_labels() |>
  italicize_levels() |>
  modify_header(estimate ~ "**OR**", std.error ~ "**SE**") |>
  modify_caption("**Table 2. Model A: Demographic Predictors of Poor Health**") |>
  as_gt()
```

::: {.callout-note}
`pattern = "{estimate}{stars}"` places the stars immediately after the
estimate. Use `"{estimate} {stars}"` to add a space. `hide_ci = TRUE` drops
the CI columns entirely.
:::

## Footnote Management

When you drop the CI columns, the default footnote defining "CI" becomes
orphaned. Remove it by filtering the internal `table_styling$abbreviation`
tibble directly.

```{r}
#| label: tbl-footnote

tbl_no_ci <- tbl_regression(
  m_B,
  exponentiate = TRUE,
  conf.int     = FALSE
) |>
  add_significance_stars(
    hide_ci  = TRUE,
    hide_se  = FALSE,
    pattern  = "{estimate}{stars}"
  ) |>
  remove_row_type(type = "reference") |>
  bold_labels() |>
  italicize_levels() |>
  modify_header(estimate ~ "**OR**", std.error ~ "**SE**")

# Surgically remove the orphaned CI footnote
tbl_no_ci$table_styling$abbreviation <-
  tbl_no_ci$table_styling$abbreviation |>
  dplyr::filter(column != "conf.low")

tbl_no_ci |>
  modify_caption("**Table 2. Model B: Demographic + SES Predictors**") |>
  as_gt()
```

## Column Name Lookup

Before customizing column headers with `modify_header()`, use
`show_header_names()` to print the exact internal column names gtsummary
assigns. This eliminates guesswork.

```{r}
#| label: show-header-names

tbl_regression(m_A, exponentiate = TRUE) |>
  show_header_names()
```

`"estimate"` = OR column, `"conf.low"` / `"conf.high"` = CI bounds,
`"std.error"` = SE. Pass these strings to `modify_header()`.

:::

# Part 3: Combining Tables

::: {.panel-tabset}

## Side by Side (`tbl_merge`)

`tbl_merge()` places multiple model tables side by side under optional
spanning headers — the standard multi-model format for journal manuscripts.

```{r}
#| label: tbl-merge

# Helper: builds a consistently formatted model table
make_tbl <- function(model) {
  tbl_regression(
    model,
    exponentiate = TRUE,
    conf.int     = FALSE
  ) |>
    add_significance_stars(
      hide_ci  = TRUE,
      hide_se  = FALSE,
      pattern  = "{estimate}{stars}"
    ) |>
    remove_row_type(type = "reference") |>
    bold_labels() |>
    italicize_levels() |>
    modify_header(estimate ~ "**OR**", std.error ~ "**SE**")
}

t_A <- make_tbl(m_A)
t_B <- make_tbl(m_B)
t_C <- make_tbl(m_C)

# Remove orphaned CI footnotes
for (tbl_obj in list(t_A, t_B, t_C)) {
  tbl_obj$table_styling$abbreviation <-
    tbl_obj$table_styling$abbreviation |>
    dplyr::filter(column != "conf.low")
}

tbl_merge(
  tbls      = list(t_A, t_B, t_C),
  tab_spanner = c(
    "**Model A: Demographic**",
    "**Model B: + SES**",
    "**Model C: Full**"
  )
) |>
  bold_labels() |>
  modify_caption("**Table 3. Logistic Regression: Predictors of Poor Self-Rated Health**") |>
  as_gt() |>
  tab_footnote(
    footnote = "OR = odds ratio; SE = standard error. Survey-weighted quasibinomial logistic regression. ***p < 0.001; **p < 0.01; *p < 0.05.",
    locations = cells_title()
  )
```

::: {.callout-tip}
Spanning headers in `tab_spanner` support markdown (`**bold**`). All
subsequent calls such as `bold_labels()` apply to the merged object.
:::

## Stacked (`tbl_stack`)

`tbl_stack()` stacks tables vertically under group headers — ideal for
comparing an unweighted model to a survey-weighted model. The difference in
confidence interval widths illustrates the design effect directly.

```{r}
#| label: tbl-stack

# Unweighted model (plain glm)
m_unwt <- glm(
  poor_health ~ us_born + genx_mil + male + race_eth15 + edu4 + famincome,
  data   = nhis_cc,
  family = binomial()
)

tbl_unwt <- tbl_regression(m_unwt, exponentiate = TRUE) |>
  remove_row_type(type = "reference") |>
  bold_labels() |>
  italicize_levels()

# Survey-weighted model (already fitted as m_B)
tbl_wt <- tbl_regression(m_B, exponentiate = TRUE) |>
  remove_row_type(type = "reference") |>
  bold_labels() |>
  italicize_levels()

tbl_stack(
  tbls         = list(tbl_unwt, tbl_wt),
  group_header = c(
    "Unweighted (plain glm)",
    "Survey-Weighted (svyglm + quasibinomial)"
  )
) |>
  modify_caption("**Table 4. Unweighted vs. Survey-Weighted Logistic Regression**") |>
  as_gt() |>
  tab_footnote(
    footnote = "Models include: nativity, birth cohort, sex, race/ethnicity, education, and family income. OR = odds ratio; 95% CI in brackets.",
    locations = cells_title()
  )
```

::: {.callout-note}
Notice how confidence intervals widen in the survey-weighted model. This
reflects the design effect: clustering and stratification reduce the
effective sample size, increasing standard errors.
:::

:::

# Part 4: Exporting to Word with flextable

For manuscript submission, most journals require `.docx` files.
`as_flex_table()` converts any gtsummary object to a flextable, which is
then exported with `save_as_docx()`.

::: {.panel-tabset}

## Convert from gtsummary

```{r}
#| label: flextable-export

ft <- tbl_merge(
  tbls      = list(t_A, t_B, t_C),
  tab_spanner = c(
    "**Model A: Demographic**",
    "**Model B: + SES**",
    "**Model C: Full**"
  )
) |>
  bold_labels() |>
  modify_caption("Table 3. Logistic Regression: Predictors of Poor Self-Rated Health") |>
  as_flex_table() |>
  set_table_properties(width = 1, layout = "autofit") |>
  fontsize(size = 10, part = "all") |>
  font(fontname = "Times New Roman", part = "all") |>
  add_footer_lines(
    "Note. OR = odds ratio; SE = standard error. Survey-weighted quasibinomial logistic regression. ***p < 0.001; **p < 0.01; *p < 0.05."
  )

ft
```

```{r}
#| label: save-docx
#| eval: false

# Always save in its own separate chunk
save_as_docx(ft, path = "output/Table3_regression.docx")
```

::: {.callout-important}
## `save_as_docx()` rule

Always call `save_as_docx()` in its **own separate chunk**. Sharing a chunk
with table-building code causes Quarto to render the table inline and write
the file simultaneously, producing duplicate output.
:::

## Standalone flextable

You can also build a flextable directly from any data frame — without going
through gtsummary — when you need full control over layout.

```{r}
#| label: flextable-standalone

summary_tbl <- nhis_cc |>
  group_by(us_born) |>
  summarise(
    N        = n(),
    Pct_poor = round(mean(poor_health) * 100, 1),
    .groups  = "drop"
  )

flextable(summary_tbl) |>
  set_header_labels(
    us_born  = "Nativity",
    N        = "N (unweighted)",
    Pct_poor = "% Poor Health"
  ) |>
  bold(part = "header") |>
  bg(bg = "#f0f0f0", part = "header") |>
  add_footer_lines("Source: NHIS teaching sample, n = 16,880 complete cases.") |>
  autofit() |>
  set_caption("Summary of Poor Health by Nativity")
```

:::

# Quick Reference

::: {.panel-tabset}

## gtsummary functions

```{r}
#| label: cheatsheet
#| echo: false

tibble::tribble(
  ~Function,                          ~Purpose,
  "`tbl_summary()`",                  "Unweighted descriptive statistics",
  "`tbl_svysummary()`",               "Survey-weighted descriptive statistics",
  "`tbl_regression()`",               "Regression table from model object",
  "`tbl_merge()`",                    "Place tables side by side",
  "`tbl_stack()`",                    "Stack tables vertically",
  "`add_overall()`",                  "Add total/overall column",
  "`add_p()`",                        "Append p-values for group comparisons",
  "`type = list(var ~ 'continuous')`","Display variable as mean (SD)",
  "`type = list(var ~ 'categorical')`","Display variable as counts (%)",
  "`statistic = list(...)`",          "Customize the displayed statistic",
  "`bold_labels()`",                  "Bold variable name rows",
  "`italicize_levels()`",             "Italicize category rows",
  "`bold_p(t = 0.05)`",               "Bold significant p-values",
  "`remove_row_type('reference')`",   "Drop reference category rows",
  "`add_significance_stars()`",       "Add * ** *** to estimates",
  "`modify_header()`",                "Rewrite column headers (use `{n}` for counts)",
  "`modify_caption()`",               "Add or change table caption",
  "`show_header_names()`",            "Print internal column name reference",
  "`as_gt()`",                        "Convert to gt for HTML rendering",
  "`as_flex_table()`",                "Convert to flextable (for Word export)",
  "`options(gtsummary.use_ftExtra = TRUE)`", "Preserve markdown formatting in Word export",
  "`set_gtsummary_theme(...)`",       "Apply a theme globally for the session"
) |>
  gt() |>
  fmt_markdown(columns = Function) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```

## Package decision guide

```{r}
#| label: pkg-comparison
#| echo: false

tibble::tribble(
  ~Task,                               ~Package,    ~Function,
  "Descriptive Table 1",               "gtsummary", "`tbl_summary()` / `tbl_svysummary()`",
  "Single regression table",           "gtsummary", "`tbl_regression()`",
  "Multiple models side by side",      "gtsummary", "`tbl_merge()`",
  "Before/after comparison (stacked)", "gtsummary", "`tbl_stack()`",
  "Word / .docx export",               "flextable", "`as_flex_table()` + `save_as_docx()`",
  "Custom standalone table",           "flextable", "`flextable()`"
) |>
  gt() |>
  fmt_markdown(columns = Function) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```

:::

---